\documentclass[pdflatex,sn-mathphys-num]{sn-jnl}


\theoremstyle{thmstyleone}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{thmstyletwo}
\newtheorem{remark}{Remark}
\theoremstyle{thmstylethree}
\newtheorem{definition}{Definition}

\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning}

% sn-mathphys-num.bst emits \burl for URLs and falls back to \textsf (which
% breaks on underscores) when \burl is undefined; route it through hyperref.
\providecommand{\burl}[1]{\url{#1}}

\graphicspath{{figures/}}

\newcommand{\vI}{\textcolor[rgb]{0.11,0.60,0.36}{\textbf{I}}}
\newcommand{\vR}{\textcolor[rgb]{0.80,0.16,0.16}{\textbf{R}}}
\newcommand{\vU}{\textcolor[rgb]{0.85,0.55,0.00}{\textbf{U}}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\Ucal}{\mathcal{U}}
\newcommand{\KL}{\mathrm{KL}}
\newcommand{\Bern}{\mathrm{Bern}}
\newcommand{\eps}{\varepsilon}
\newcommand{\Efrak}{\mathbb{E}}

\raggedbottom

\begin{document}

\title[Anytime-Valid Certificates for Machine Unlearning]{VOUCH: Anytime-Valid Certificates for Machine Unlearning via Paired-Canary Hypothesis-Betting}

\author*[1]{\fnm{Quang-Vinh} \sur{Dang}}\email{vinh.dq4@buv.edu.vn}

\affil*[1]{\orgname{British University Vietnam}, \orgaddress{\city{Hung Yen}, \country{Vietnam}}}

\abstract{Machine unlearning methods for large language models are deployed to honour
deletion requests, yet they ship with no statistical certificate that forgetting
actually occurred. The evaluations in common use are fixed-sample tests: they are
unsound as membership evidence unless what enters training is randomised, their error
guarantees collapse the moment an auditor looks at the evidence as it accumulates, and
they lean on a retrained reference model that is unaffordable at scale and, for
multi-stage pipelines, not even well defined. We introduce \emph{VOUCH}, an
anytime-valid statistical certification framework that removes all three obstacles at
once. At fine-tuning time the provider plants \emph{paired ghost canaries}:
exchangeable twin records of which a fair coin admits one into training (and later into
the forget set, so that it traverses the entire unlearning pipeline) while the other is
never shown to any model. Because exact unlearning leaves the two twins statistically
interchangeable, the within-pair score difference is symmetric \emph{by construction};
this hands the verifier an exact, distribution-free, finite-sample null that needs
neither shadow models nor a retrained reference nor any assumption about the model.
Over this null VOUCH runs betting e-processes that remain valid at every stopping time
simultaneously, and from them derives three coupled objects: an anytime-valid confidence
sequence on the residual membership advantage, an $(\eps,\alpha)$-forgetting certificate
obtained by sequential two-sided equivalence testing, and a revocation alarm. We prove
that valid composition runs in opposite directions for the two claims---certificates
compose by consensus, alarms by product---and show empirically that the intuitive
alternatives are invalid, falsely certifying up to $99.6\%$ of the time. Across a
two-thousand-seed calibration study under adversarial optional stopping and end-to-end
pipelines on the TOFU and MUSE-News benchmarks over several architectures, VOUCH
certifies genuine unlearning, revokes un-unlearned models within a handful of canary
pairs, and---by pairing every certificate with a utility reading---exposes
gradient-ascent unlearning that forgets by destroying the model. Verification costs on
the order of $10^4$ forward passes, roughly two orders of magnitude below shadow-model
evaluation.}

\keywords{machine unlearning, anytime-valid inference, e-values, membership inference,
certification, large language models}

\maketitle

\section{Introduction}\label{sec:intro}

When a user invokes the right to erasure, a model provider must do more than run an
unlearning routine: it must be able to \emph{show}, to a regulator or an internal audit
function, that the routine worked. This obligation is now concrete. The EU AI Act's
audit provisions, Article~17 of the GDPR, and a growing set of national data-protection
regimes all presume that erasure can be demonstrated, not merely asserted. The
algorithmic side of machine unlearning has risen to meet the demand: gradient ascent
and its stabilised successors, negative preference optimisation
\citep{zhang2024npo,fan2025simnpo}, representation misdirection \citep{li2024wmdp},
and task-vector negation now remove targeted knowledge from large models
at modest cost. The \emph{verification} side has not kept pace, and it is verification,
not optimisation, that a regulator cares about.

Three obstacles stand between current practice and a defensible certificate, and it is
worth stating them plainly because VOUCH is organised around dismantling each one.
\emph{First}, a high membership-inference score on a supposedly forgotten example is
not evidence that the example was ever trained on. Distribution shift and the model's
own priors confound the signal, and \citet{zhang2024mia} have shown formally that
membership cannot be established without controlling the randomness of what enters
training. \emph{Second}, deletion requests arrive as a stream and auditors inspect the
evidence as it accrues, stopping when they are satisfied; every fixed-sample test in
current use---the forget-quality metric of \citet{maini2024tofu}, the leakage metric of
\citet{shi2024muse}, the shadow-model panels of \citet{hayes2024inexact}---has its
error guarantee voided by exactly this optional stopping, and we measure the damage in
Section~\ref{sec:exp-validity}. \emph{Third}, the comparators these tests rely on---a
model retrained from scratch without the deleted data---are prohibitively expensive at
the scale of modern language models and, as \citet{yu2025impossibility} show, are not
even well defined for the multi-stage pipelines used in practice, because
path-dependence means no unique retrained model exists to compare against.

VOUCH dissolves all three obstacles with a single design decision and a single
inferential engine. The design decision is the \emph{paired ghost canary}. At
fine-tuning time we generate exchangeable twin records, and for each pair we flip a
fair coin to decide which twin is admitted into training; the admitted twin is later
appended to the forget set, so that it passes through the very same unlearning pipeline
as organic data, while its twin is withheld from every model. This construction is what
makes the audit sound: under exact unlearning the model is independent of the coin, so
the two twins are exchangeable and the difference between their scores is symmetric
about zero \emph{regardless of the model, the scoring function, or the data
distribution}. We therefore obtain an exact, distribution-free, finite-sample null that
we generated ourselves, rather than one we must assume. The engine is the betting
e-process from game-theoretic statistics, which turns this null into a test whose
validity holds not at one predetermined sample size but at every stopping time at once,
by Ville's inequality. Streaming deletions and continuous public monitoring, fatal to a
fixed-sample test, are precisely what an e-process is built to survive.

Building the framework and running it end-to-end forced four corrections and
refinements on the initial design, and because each turned out to matter in practice we
foreground them rather than bury them. We found that the natural way to combine several
attack scores into one certificate---bet on an adaptively weighted mixture of their
signs---is \emph{anytime-invalid}, and in simulation it falsely certified a leaking
model $99.6\%$ of the time; the sound alternative requires every score's own process to
concur. We found that certificates and alarms must compose in \emph{opposite}
directions across deletion waves, and that multiplying certificate e-values, the
instinctive move, certifies histories that contain a genuinely bad wave. We found on
real language models that gradient-ascent unlearning frequently \emph{over-forgets},
driving the forgotten twins to scores conspicuously \emph{below} their never-seen
partners---membership leakage that a one-sided test happily certifies and that a
two-sided test must catch. And we found, when we corrected the sample-size analysis
against the information-theoretic lower bound, that an earlier back-of-envelope estimate
understated the required cohort by threefold. Each of these is now a theorem or a
measured result rather than a footnote.

The contributions of this paper are as follows. We formulate unlearning certification
as sequential two-sided equivalence testing against a declared attack class, with an
exact distribution-free null supplied by paired ghost canaries
(Section~\ref{sec:framework}). We give the supporting theory: exact validity at any
stopping time, sound composition across scores and across deletion waves in both
directions, power that matches the Kullback--Leibler information limit, a
magnitude-aware revocation arm that is exact under the symmetry null, and a bridge from
certified $(\eps_u,\delta_u)$-unlearning (Section~\ref{sec:theory}). Finally we report a
comprehensive empirical study (Section~\ref{sec:exp}): calibration under adversarial
optional stopping over two thousand seeds; end-to-end certification of gradient ascent,
GradDiff, negative preference optimisation, and retraining-from-scratch on the TOFU and
MUSE-News benchmarks across two model architectures, with utility guardrails and
dose--response calibration; recoverability probes that relearn, quantise, and jailbreak
the unlearned model; a model axis spanning 2019 to 2026; and cost measurements showing
that a certificate takes seventy-six seconds on a single mid-range GPU. All code, all
results, and the fault-tolerant harness that produced them on free ephemeral compute are
released.

\section{Related work}\label{sec:related}
\input{sections/related}

\section{Setting and the VOUCH protocol}\label{sec:framework}

This section develops the protocol in full: the actors and threat model, the canary
construction that manufactures the null, the certified quantity, the betting engine
with its explicit wealth dynamics, and finally the assembled protocol as pseudocode
(Algorithm~\ref{alg:vouch}) and as a flowchart (Figure~\ref{fig:flowchart}).

\subsection{Actors, pipeline, and threat model}\label{sec:setting}

The provider fine-tunes a base model on
$D_{\text{keep}} \cup D_{\text{forget}} \cup C$, where $C$ is the canary set described
below. When deletion requests arrive, an unlearning algorithm $\Ucal$---negative
preference optimisation, GradDiff, and so on---produces an unlearned model $M_u$ from
the fine-tuned model, using the forget set augmented with the relevant canaries. A
verifier, holding only black-box query access to $M_u$ together with a canary manifest
whose hash was published before unlearning began, runs VOUCH and either issues,
withholds, or revokes a certificate. Throughout we assume the honest-but-verifiable
setting: the provider genuinely executes $\Ucal$ and the open question is whether it
worked and how well, under quantified error control. A provider willing to special-case
the committed canaries is outside our scope and, as prior work establishes, can only be
addressed by cryptographic attestation, with which VOUCH composes.

\subsection{Paired ghost canaries}\label{sec:canaries}

A templated generator emits exchangeable twin pairs
$(c_i^0, c_i^1)$ for $i = 1,\dots,m$. The two twins in a pair are independent draws from
the same conditional template distribution---the same surface form populated with
independently sampled fictitious entities carrying secrets of at least forty bits of
entropy---so that within a pair they are exchangeable by construction. We match the
template to the host corpus, using question--answer pairs for TOFU and record-style
sentences for news text, so that the canaries do not stand out. For each pair an
independent fair coin $b_i \sim \Bern(\tfrac12)$ selects the \emph{in-twin}
$c_i^{b_i}$, which we insert into
the fine-tuning corpus with a repetition factor $r_i \in \{1,2,4,8\}$ (these strata give
us a dose--response curve) and later append to the forget set, so that it traverses the
unlearning pipeline exactly as organic forget data does. The \emph{ghost twin}
$c_i^{1-b_i}$ is never
shown to any model. Before unlearning begins the provider publishes the commitment
\begin{equation}\label{eq:commit}
\mathrm{commit} \;=\; H\bigl(\{(c_i^0,\, c_i^1,\, b_i,\, r_i)\}_{i=1}^{m}\bigr),
\qquad H = \text{SHA-256},
\end{equation}
which binds the pairs and the coins and prevents the audit from being gamed after the
fact: at verification time the manifest is revealed and checked against
\eqref{eq:commit} before a single query is made.

\subsection{Scores and the certified quantity}\label{sec:scores}

We fix a declared class $\F$ of
scalar score functions---the token-normalised negative log-likelihood of the secret
span, the min-$k\%$ token log-probability, the base-model-calibrated loss ratio, and
optionally a linear probe on the hidden state---each aggregated over $Q$ fixed query
prompts. For pair $i$ and score $s$ we form the within-pair difference and its sign,
\begin{equation}\label{eq:sign}
D_i^{(s)} = s(M_u, c_i^{\text{in}}) - s(M_u, c_i^{\text{ghost}}), \qquad
Z_i^{(s)} = \mathbf{1}\{D_i^{(s)} > 0\},
\end{equation}
breaking ties by a committed fair coin. Writing $p^{(s)} = \Pr[Z^{(s)} = 1]$ and
$\Delta^{(s)} = 2p^{(s)} - 1$, the quantity we certify is $\sup_{s\in\F}|\Delta^{(s)}|$,
the largest residual advantage any score in the class can extract. We are deliberate
that this is influence extractable \emph{through $\F$} on the canary population: it is
the strongest target compatible with the impossibility results for behavioural audits,
and we make no claim about erasure in parameter space. The certificate itself is worth
pinning down precisely, since everything that follows is engineered to earn it.

\begin{definition}[$(\eps,\alpha)$-forgetting certificate]\label{def:cert}
Fix a tolerance $\eps \in (0,1)$, an error budget $\alpha \in (0,1)$, and the boundary
$p_0 = \tfrac12 + \tfrac\eps2$. An $(\eps,\alpha)$-forgetting certificate for $M_u$
over the class $\F$ is a data-dependent stopping decision that asserts
$\sup_{s\in\F}|\Delta^{(s)}| < \eps$, issued by a procedure with the guarantee that,
whenever in truth $\sup_{s\in\F}|\Delta^{(s)}| \ge \eps$, the probability that a
certificate is \emph{ever} issued---at any stopping time, under any monitoring rule---is
at most $\alpha$.
\end{definition}

The two-sidedness in Definition~\ref{def:cert} is deliberate: $\Delta^{(s)} \le -\eps$,
the over-forgetting regime in which the in-twin scores conspicuously \emph{below} its
never-seen partner, is membership leakage exactly as much as $\Delta^{(s)} \ge \eps$
is, and Section~\ref{sec:exp} shows that gradient-ascent unlearning lands there in
practice.

\subsection{The betting engine}\label{sec:engine}

All three of VOUCH's outputs are built from one primitive, the \emph{betting
e-process}, and it is worth writing its dynamics out explicitly because the validity
proofs in Section~\ref{sec:theory} reduce to inspecting them. A gambler starts with
unit wealth and, before pair $i$ is revealed, stakes a predictable fraction
$\lambda_i$ (chosen from the information $\F_{i-1}$ revealed so far) on the sign
$Z_i$ deviating from a hypothesised mean. Against the null $H_0^{+}\colon
\Efrak[Z] \ge p_0$ (``the in-twins still score high'') and its mirror
$H_0^{-}\colon \Efrak[Z] \le 1-p_0$, the wealth processes for score $s$ are
\begin{equation}\label{eq:wealth}
E_t^{s,+} \;=\; \prod_{i \le t}\bigl(1 + \lambda_i^{s,+}\,(p_0 - Z_i^{(s)})\bigr),
\qquad
E_t^{s,-} \;=\; \prod_{i \le t}\bigl(1 + \lambda_i^{s,-}\,(Z_i^{(s)} - (1 - p_0))\bigr),
\end{equation}
with stakes confined to $\lambda_i^{s,\pm} \in [0,\, 1/(1-p_0))$ so that wealth can
never go negative. Under any distribution in the respective null each factor has
conditional mean at most one, so each process is a nonnegative supermartingale with
initial value one, and Ville's inequality converts wealth into evidence: the chance
that a supermartingale ever exceeds $1/\alpha$ is at most $\alpha$. Wealth above
$1/\alpha$ is therefore a licence to reject that null at whatever time we happen to be
looking---the anytime validity on which the whole framework rests.

How the stakes are chosen determines power, not validity, which frees us to be
aggressive. VOUCH plays a uniform mixture of eight experts and lets their wealth
arbitrate: six constant-fraction bettors $\lambda = c \cdot \lambda_{\max}$ with
$c \in \{0.02, 0.05, 0.1, 0.2, 0.4, 0.7\}$, an online-Newton-step (ONS) learner, and a
truncated Krichevsky--Trofimov (KT) plug-in. The ONS expert follows the coin-betting
recursion
\begin{equation}\label{eq:ons}
g_i = \frac{p_0 - Z_i}{1 + \lambda_i (p_0 - Z_i)}, \qquad
A_i = A_{i-1} + g_i^2, \qquad
\lambda_{i+1} = \Pi_{[0,\lambda_{\max}]}\!\Bigl(\lambda_i + \tfrac{\eta\, g_i}{A_i}\Bigr),
\end{equation}
with $A_0 = 1$, $\eta = \tfrac12$, and $\Pi$ the clip onto the admissible range; its
log-wealth regret against the best constant bet in hindsight is $O(\log t)$. The KT
expert bypasses stake selection altogether: it estimates the mean by
$q_i = (\sum_{j<i} Z_j + \tfrac12)/i$, truncates $q_i$ to the alternative side of
$p_0$, and multiplies wealth by the plug-in likelihood ratio
\begin{equation}\label{eq:kt}
e_i \;=\; Z_i\,\frac{q_i}{p_0} \;+\; (1 - Z_i)\,\frac{1 - q_i}{1 - p_0},
\end{equation}
which has conditional mean at most one under the null (it is linear in the mean, equals
one at $p_0$, and decreases into the null) and achieves the optimal growth rate up to a
$\tfrac1{2t}\log t$ redundancy. Because a convex combination of e-processes with a
fixed prior is again an e-process, the mixture inherits exact validity while paying
only a $\log 8$ additive cost in log-wealth relative to its best expert---power adapts
online with nothing to tune.

The same wealth idea, run over a grid rather than at a single boundary, yields the
quantitative face of the certificate. For every candidate mean $m$ on a grid of
$1001$ points in $[0,1]$ we maintain two wealth processes betting that the truth is
above and below $m$, with empirical-Bernstein (aGRAPA) stakes
\begin{equation}\label{eq:agrapa}
\lambda_i(m) \;=\; \Pi_{[0,\,\lambda_{\max}(m)]}
\Bigl(\tfrac{\hat\mu_{i-1} - m}{\hat\sigma_{i-1}^2 + (\hat\mu_{i-1} - m)^2}\Bigr),
\end{equation}
where $\hat\mu_{i-1}$ and $\hat\sigma^2_{i-1}$ are the running mean and variance of the
signs. The set of grid values whose wealth on both sides stays below $2/\alpha$,
intersected over time,
\begin{equation}\label{eq:cs}
C_t \;=\; \bigcap_{u \le t}\,\bigl\{\, m :\ \max\bigl(W_u^{\uparrow}(m),\,
W_u^{\downarrow}(m)\bigr) < 2/\alpha \,\bigr\},
\end{equation}
is a $(1-\alpha)$ confidence sequence for $p^{(s)}$: it covers the truth at every time
simultaneously, so the auditor may glance at it whenever they please. Mapped through
$\Delta = 2p - 1$ it reports the residual advantage with its anytime-valid uncertainty.

The revocation arm inverts the role of the null. Under \emph{exact} unlearning,
Theorem~\ref{thm:null} makes every $Z_i^{(s)}$ an exact fair coin, so here---and only
here---adaptive score weighting is sound. We run, per score and per direction,
e-processes against $\Efrak[Z] = \tfrac12$ and combine them with predictable
wealth-proportional weights: with $w_{s,i} \propto \exp(\log E^{s}_{i-1})$ computed
before pair $i$ is revealed,
\begin{equation}\label{eq:revmix}
R_t \;=\; \prod_{i\le t}\;\sum_{s\in\F} w_{s,i}\,
\frac{E^{s}_{i}}{E^{s}_{i-1}},
\end{equation}
which is an e-process because each factor is a predictable convex combination of
conditional-mean-one increments. When $R_t \ge 1/\alpha$ the verifier holds
anytime-valid evidence of residual memorisation and the certificate is withheld or
revoked; the weights meanwhile steer wealth toward whichever score discriminates best,
which is exactly what an alarm should do.

\subsection{The protocol end to end}\label{sec:protocol}

Algorithm~\ref{alg:vouch} assembles the pieces into the three-phase protocol, and
Figure~\ref{fig:flowchart} shows the same pipeline as a flowchart. Phase~0 runs at
design time: generate the cohort, flip the coins, publish the commitment, fine-tune
with the in-twins planted. Phase~1 is the provider's unlearning step, in which the
in-twins ride along with the organic forget set. Phase~2 is the audit: reveal and check
the manifest, then stream the pairs in random order, updating the certificate
processes, the confidence sequences, and the revocation mixture after each pair, and
act on whichever boundary is crossed first. R-VOUCH then re-runs the same loop on
probed variants of $M_u$---after relearning on public data, after 4-bit quantisation,
and under adversarial prompt wrappers---and the robust certificate requires every probe
to pass. Streaming deletion waves each draw a fresh cohort with independent coins;
Theorem~\ref{thm:streaming} dictates how their verdicts may be combined.

\begin{algorithm}[t]
\caption{The VOUCH protocol}\label{alg:vouch}
\begin{algorithmic}[1]
\Statex \textbf{Phase 0 (design time, before fine-tuning)}
\State generate twin pairs $\{(c_i^0, c_i^1)\}_{i=1}^m$; draw coins
       $b_i \sim \Bern(\tfrac12)$ and strata $r_i \in \{1,2,4,8\}$
\State $\textsc{In} \gets \{c_i^{b_i}\}$;\quad $\textsc{Ghost} \gets \{c_i^{1-b_i}\}$
       \Comment{ghosts are never shown to any model}
\State publish $\mathrm{commit} = H(\{(c_i^0, c_i^1, b_i, r_i)\})$
       \Comment{SHA-256, Eq.~\eqref{eq:commit}}
\State fine-tune $M_{\mathrm{ft}}$ on $D_{\text{keep}} \cup D_{\text{forget}} \cup
       (\textsc{In} \text{ with repetitions } r_i)$
\Statex \textbf{Phase 1 (unlearning)}
\State $M_u \gets \Ucal(M_{\mathrm{ft}};\ D_{\text{forget}} \cup \textsc{In})$
       \Comment{in-twins traverse the pipeline as organic forget data}
\Statex \textbf{Phase 2 (certification; black-box queries only)}
\State reveal manifest; \textbf{abort} unless $H(\text{manifest}) = \mathrm{commit}$
\State initialise $E^{s,+} = E^{s,-} = 1$ for all $s \in \F$; revocation mixture
       $R = 1$; CS grids $C^{(s)}$
\For{$i$ in a committed random order of $1,\dots,m$}
  \For{$s \in \F$}
     \State $D_i^{(s)} \gets s(M_u, c_i^{\text{in}}) - s(M_u, c_i^{\text{ghost}})$;
            \quad $Z_i^{(s)} \gets \mathbf{1}\{D_i^{(s)} > 0\}$ (ties: committed coin)
     \State update $E^{s,+}, E^{s,-}$ by Eq.~\eqref{eq:wealth} with mixture bets
            \eqref{eq:ons}--\eqref{eq:kt}; update $C^{(s)}$ by
            \eqref{eq:agrapa}--\eqref{eq:cs}
  \EndFor
  \State update the revocation mixture $R$ by Eq.~\eqref{eq:revmix}
  \If{$R \ge 1/\alpha$} \textbf{revoke}; \Return
        $\textsc{Fail}(\text{evidence} = R)$
  \EndIf
  \If{$\min_{s,\pm} E^{s,\pm} \ge 1/\alpha$} \textbf{break}
        \Comment{certificate earned early}
  \EndIf
\EndFor
\State run R-VOUCH probes P1 (relearn), P2 (quantise), P3 (jailbreak); repeat the
       loop above per probe
\If{$\min_{s,\pm} E^{s,\pm} \ge 1/\alpha$ \textbf{and} every probe passes}
   \State \textbf{issue} $\mathrm{cert} = (\eps, \alpha, t, C_t, \mathrm{commit},
          \F, \text{probe report})$
\Else
   \State \textbf{withhold} \Comment{undetermined: report $C_t$ and continue
          monitoring}
\EndIf
\Statex \textbf{Streaming (waves $k = 1, 2, \dots$)}: fresh cohort and coins per wave;
        global certificate $=$ all-pass; global alarm $=$ $\prod_k R^{(k)}$; per-wave
        alarms at $\alpha_k = \alpha 2^{-(k+1)}$ \Comment{Theorem~\ref{thm:streaming}}
\end{algorithmic}
\end{algorithm}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  font=\footnotesize,
  node distance=4mm and 6mm,
  box/.style={draw, rounded corners=2pt, align=center, inner sep=3.5pt,
              minimum height=7mm},
  data/.style={box, fill=gray!12},
  proc/.style={box, fill=blue!9},
  eng/.style={box, fill=violet!10},
  good/.style={box, fill=green!18},
  bad/.style={box, fill=red!16},
  wait/.style={box, fill=orange!15},
  arr/.style={-{Latex[length=1.8mm]}, semithick},
  lab/.style={font=\scriptsize\itshape, text=black!60}
]
% Phase 0 row
\node[proc] (gen) {twin pairs $(c_i^0,c_i^1)$\\ strata $r_i\in\{1,2,4,8\}$};
\node[proc, right=of gen] (coin) {fair coins\\ $b_i\sim\Bern(\frac12)$};
\node[data, right=of coin] (split) {$\textsc{In}=\{c_i^{b_i}\}$\\
  $\textsc{Ghost}=\{c_i^{1-b_i}\}$};
\node[data, right=of split] (commit) {publish\\ $H(\text{manifest})$};
\node[proc, below=6mm of split] (ft) {fine-tune $M_{\mathrm{ft}}$ on
  $D_{\text{keep}}\cup D_{\text{forget}}\cup \textsc{In}$};
\node[proc, below=of ft] (unlearn) {unlearn: $M_u \gets \Ucal(M_{\mathrm{ft}};\,
  D_{\text{forget}}\cup\textsc{In})$};
\node[proc, below=of unlearn] (score) {reveal manifest, check hash;\\
  stream pairs: $D_i^{(s)},\ Z_i^{(s)}$ by Eq.~\eqref{eq:sign}};
% three coupled processes
\node[eng, below=7mm of score] (cs) {confidence sequence\\
  $C_t$ on each $\Delta^{(s)}$, Eq.~\eqref{eq:cs}};
\node[eng, left=5mm of cs] (cert) {certificate arm\\
  $E_t^{s,\pm}$, Eq.~\eqref{eq:wealth}\\ (min over $s,\pm$)};
\node[eng, right=5mm of cs] (rev) {revocation arm\\
  mixture $R_t$, Eq.~\eqref{eq:revmix}};
% decisions
\node[good, below=7mm of cert, text width=30mm] (issue)
  {all $E^{s,\pm}\!\ge\!\frac1\alpha$, probes pass: \textbf{issue}
   $(\eps,\alpha)$-cert.};
\node[bad, below=7mm of rev, text width=24mm] (revoke) {$R_t \ge \frac1\alpha$:\\
  \textbf{revoke / withhold}};
\node[wait, below=7mm of cs, text width=24mm] (undet) {otherwise: report $C_t$,
  keep streaming};
\node[eng, below=4mm of undet] (probes) {R-VOUCH probes: relearn, quantise,
  jailbreak $\to$ rerun loop};
% arrows
\draw[arr] (gen) -- (coin);
\draw[arr] (coin) -- (split);
\draw[arr] (split) -- (commit);
\draw[arr] (split) -- (ft);
\draw[arr] (ft) -- (unlearn);
\draw[arr] (unlearn) -- (score);
\draw[arr] (commit.south) to[out=-90, in=20] (score.east);
\draw[arr] (score) -- (cert);
\draw[arr] (score) -- (cs);
\draw[arr] (score) -- (rev);
\draw[arr] (cert) -- (issue);
\draw[arr] (rev) -- (revoke);
\draw[arr] (cs) -- (undet);
\draw[arr] (undet) -- (probes);
\draw[arr] (probes.west) to[out=180, in=180, looseness=1.4] (cert.west);
% phase labels
\node[lab, left=2mm of gen.west, anchor=east] {Phase 0};
\node[lab, left=48mm of unlearn.west, anchor=east] {Phase 1};
\node[lab, left=41mm of score.west, anchor=east] {Phase 2};
\end{tikzpicture}
\caption{The VOUCH pipeline. Phase 0 plants paired ghost canaries and commits to the
manifest; Phase 1 unlearns with the in-twins riding along; Phase 2 streams the pairs
through three coupled anytime-valid processes---the per-score certificate minimum, the
betting confidence sequence, and the revocation mixture---whose boundary crossings
trigger issuance, continued monitoring, or revocation. R-VOUCH probes rerun the loop
on relearned, quantised, and jailbroken variants before a robust certificate is
issued.}
\label{fig:flowchart}
\end{figure}

\section{Theory}\label{sec:theory}

Our guarantees rest on one structural fact---the null we plant is exact---and on the
supermartingale property of betting. The proofs are short, and their brevity is the
point: the design does the work, so we give every proof in full. Throughout,
$\F_{i-1}$ denotes the verifier's filtration, generated by everything revealed through
pair $i-1$; all bets and mixture weights are $\F_{i-1}$-measurable (predictable).

\begin{theorem}[Exact, distribution-free null]\label{thm:null}
Suppose the unlearned model $M_u$ is conditionally independent of the coin $b_i$ given
the pair $(c_i^0, c_i^1)$; exact unlearning and retraining-from-scratch both satisfy
this. Then within-pair exchangeability gives $D_i^{(s)} \overset{d}{=} -D_i^{(s)}$, so
$Z_i^{(s)} \sim \Bern(\tfrac12)$ exactly---for any model, any score, and any set of
query prompts.
\end{theorem}

\begin{proof}
Condition on the unordered pair $\{c_i^0, c_i^1\}$ and the past $\F_{i-1}$. The twins
are exchangeable within the pair by construction, and under the hypothesis $M_u$ is
conditionally independent of $b_i$: the pipeline either saw neither twin or saw only
the in-twin, whose identity is determined by $b_i$, and after exact unlearning the
output law no longer depends on that identity. The joint law of
$(c_i^{\text{in}}, c_i^{\text{ghost}}, M_u)$ is therefore invariant under swapping the
two twins. Swapping negates
$D_i^{(s)} = s(M_u, c_i^{\text{in}}) - s(M_u, c_i^{\text{ghost}})$ while preserving its
distribution, so $D_i^{(s)} \overset{d}{=} -D_i^{(s)}$ conditionally on $\F_{i-1}$.
Symmetry of $D_i^{(s)}$ gives $\Pr[D_i^{(s)} > 0 \mid \F_{i-1}] =
\Pr[D_i^{(s)} < 0 \mid \F_{i-1}]$, and the committed tie-break coin distributes any
atom at zero equally between the two signs, so
$Z_i^{(s)} \mid \F_{i-1} \sim \Bern(\tfrac12)$ exactly. No property of the model, the
score, or the query set was used beyond measurability.
\end{proof}

Theorem~\ref{thm:null} is the reason VOUCH needs neither shadow models nor a retrained
reference: the null distribution is not estimated or assumed but generated by our own
coins, and it is exact at finite $m$. Every process below inherits its validity from it.

\begin{theorem}[Certificate coverage and sound multi-score composition]\label{thm:cert}
Issue the certificate only when \emph{every} per-score, per-direction e-process exceeds
$1/\alpha$. Then for any class $\F$ and any stopping rule,
$\Pr[\text{a false }(\eps,\alpha)\text{-certificate is ever issued}] \le \alpha$
whenever $\sup_s |\Delta^{(s)}| \ge \eps$. No multiplicity correction is required,
because issuance forces the truly-violating score's own supermartingale across
$1/\alpha$, an event of probability at most $\alpha$ by Ville's inequality.
\end{theorem}

\begin{proof}
Suppose the certification claim is false: some score $s^\ast$ and direction violate the
tolerance, say $p^{(s^\ast)} \ge p_0$ (the case $p^{(s^\ast)} \le 1 - p_0$ is
symmetric). Consider that score's own upper process $E^{s^\ast,+}$ from
Eq.~\eqref{eq:wealth}. Each of its factors has conditional mean
\[
\Efrak\bigl[\,1 + \lambda_i (p_0 - Z_i^{(s^\ast)}) \mid \F_{i-1}\bigr]
= 1 + \lambda_i \bigl(p_0 - \Efrak[Z_i^{(s^\ast)} \mid \F_{i-1}]\bigr) \;\le\; 1,
\]
since $\lambda_i \ge 0$ is predictable and the conditional mean of the sign is at
least $p_0$. Hence $E^{s^\ast,+}$ is a nonnegative supermartingale with
$E^{s^\ast,+}_0 = 1$, and Ville's inequality gives
$\Pr[\exists t: E_t^{s^\ast,+} \ge 1/\alpha] \le \alpha$. Issuing the certificate
requires \emph{every} process to exceed $1/\alpha$---in particular the violating
score's own---so a false certificate is issued with probability at most $\alpha$,
uniformly over stopping rules. No union bound over $\F$ is needed because the argument
runs entirely through the single process attached to the violating score.
\end{proof}

\begin{remark}[Why the intuitive mixture fails]\label{rem:mixture}
It is tempting to run a \emph{single} certificate process on the adaptively reweighted
sign $\bar Z_i = \sum_s w_s Z_i^{(s)}$, and indeed this is valid for the revocation arm,
whose null makes every component an exact fair coin. It is \emph{not} valid for the
certificate, whose composite null is ``some score still leaks.'' The failure is easy to
exhibit: take $p^{(s_1)} = p_0$ exactly at the boundary and
$p^{(s_2)} = \dots = \tfrac12$ clean. Then
$\Efrak[\bar Z_i \mid \F_{i-1}] = w_1 p_0 + (1 - w_1)/2 < p_0$ whenever $w_1 < 1$, so
each factor of the mixture's certificate process has conditional mean strictly greater
than one: the process is a \emph{sub}martingale under the null and crosses $1/\alpha$
with probability approaching one. Worse, the violation is generic rather than
adversarial, because discrimination-driven reweighting moves weight \emph{away} from
the one leaking score under partial unlearning. We measured a false-certification rate
of $0.996$ for this construction against $0.007$ for the per-score minimum
(Table~\ref{tab:soundness}), which is why VOUCH requires consensus rather than an
average.
\end{remark}

\begin{theorem}[Power]\label{thm:power}
If $|\Delta^{(s)}| < \eps$ for all $s$ and the bets have sublinear regret (the mixture
of Section~\ref{sec:engine} qualifies), then
$\tfrac1t \log E_t \to \KL(\Bern(p)\,\|\,\Bern(p_0))$ almost
surely for each process, and the expected certification time is
$\Efrak[\tau^*] = (1+o(1))\,\log(1/\alpha)/\min_{s,\pm}\KL$. For exact unlearning and
small $\eps$ this is approximately $8\log(1/\alpha)/\eps^2$ pairs.
\end{theorem}

\begin{proof}
Fix one process, say the upper process for a score with true sign probability
$p < p_0$. For a constant bet $\lambda$, the log-wealth increments
$\log(1 + \lambda(p_0 - Z_i))$ are i.i.d.\ with mean
$f(\lambda) = p \log(1 - \lambda(1 - p_0)) + (1-p)\log(1 + \lambda p_0)$, so by the
strong law $\tfrac1t \log E_t \to f(\lambda)$ almost surely. The Kelly-optimal
constant bet is $\lambda^\ast = (p_0 - p)/(p_0(1 - p_0))$: substituting it turns the
two possible wealth factors into the Bernoulli likelihood ratios
\[
1 - \lambda^\ast(1 - p_0) = \frac{p}{p_0}, \qquad
1 + \lambda^\ast p_0 = \frac{1 - p}{1 - p_0},
\]
so $f(\lambda^\ast) = p \log\tfrac{p}{p_0} + (1-p)\log\tfrac{1-p}{1-p_0}
= \KL(\Bern(p)\,\|\,\Bern(p_0))$, and no constant bet can do better because $f$ is
concave with this maximum. A betting strategy whose log-wealth regret against the best
constant bet is $o(t)$ therefore attains the same limit; the ONS expert has $O(\log t)$
regret, the KT plug-in \eqref{eq:kt} achieves
$\KL - \tfrac{1}{2t}\log t$ directly by the Krichevsky--Trofimov redundancy bound, and
the uniform mixture over eight experts trails its best expert by at most the constant
$\log 8$, so all three satisfy the hypothesis. For the crossing time, the process
rejects when $\log E_t \ge \log(1/\alpha)$; since $\tfrac1t\log E_t \to \KL$ almost
surely, the first-crossing time $\tau^\ast$ of the drifting random walk satisfies
$\Efrak[\tau^\ast] = (1 + o(1)) \log(1/\alpha)/\KL$ by the standard Wald argument, and
issuance waits for the \emph{slowest} process, whose drift is $\min_{s,\pm} \KL$.
Finally, under exact unlearning $p = \tfrac12$ for every score, and a second-order
expansion gives $\KL(\Bern(\tfrac12) \| \Bern(\tfrac12 + \tfrac\eps2)) =
\tfrac{\eps^2}{8} + O(\eps^4)$, whence
$\Efrak[\tau^\ast] \approx 8\log(1/\alpha)/\eps^2$.
\end{proof}

Theorem~\ref{thm:power} matters for deployment because it tells the provider how large a
canary cohort to plant, and it corrects an earlier heuristic: an $\eps = 0.1$
certificate at $\alpha = 0.05$ requires on the order of six hundred pairs, not the
one-hundred-ninety a naive calculation suggested, and because the drift equals the
Kullback--Leibler divergence---the information-theoretic optimum for testing
$\Bern(\tfrac12)$ against $\Bern(p_0)$---no valid procedure can do better by more than
a constant factor.

\begin{theorem}[Streaming composition, in both directions]\label{thm:streaming}
Let deletion waves $k = 1,\dots,K$ use cohorts with independent coins.
\emph{(a) Certificates compose by consensus.} Claiming that the whole history is clean
only when every wave has individually earned its certificate keeps the probability of a
false history-wide claim at or below $\alpha$ for any $K$; multiplying certificate
e-values instead tests the intersection null ``every wave is bad,'' whose rejection
certifies merely that \emph{some} wave is clean.
\emph{(b) Alarms compose by product.} Under the intersection null ``every wave was
exactly unlearned'' each per-wave revocation process is a supermartingale, so their
product is one too and crosses $1/\alpha$ with probability at most $\alpha$; the product
accumulates leakage spread so thinly across waves that no single wave reveals it.
\emph{(c)} Per-wave alarms with $\alpha$-spending $\alpha_k = \alpha 2^{-(k+1)}$ control
family-wise false revocation over an unbounded history.
\end{theorem}

\begin{proof}
\emph{(a)} Suppose the history-wide claim is false, so some wave $k^\ast$ truly has
$\sup_s |\Delta^{(s),(k^\ast)}| \ge \eps$. The all-pass rule claims only when every
wave has individually issued, so claiming requires wave $k^\ast$'s own per-score
minimum process to cross $1/\alpha$, an event of probability at most $\alpha$ by
Theorem~\ref{thm:cert}---independently of $K$ and of how the auditor schedules the
waves. For the negative half of the statement, note that a product
$\prod_k E^{(k)}$ of \emph{certificate} e-values is, by construction, an e-value for
the intersection null $\bigcap_k \{\text{wave } k \text{ bad}\}$; rejecting an
intersection establishes only the disjunction ``some wave is clean.'' In the
one-bad-wave scenario the clean waves' e-values grow without bound
(Theorem~\ref{thm:power}) and drag the product over $1/\alpha$ regardless of the bad
wave; we measured this as a $0.96$ false history-wide certification rate against
$0.000$ for all-pass (Table~\ref{tab:streaming}).

\emph{(b)} Under the intersection null, Theorem~\ref{thm:null} applies wave by wave,
so each $R^{(k)}$ is a nonnegative supermartingale with initial value one with respect
to the combined filtration, and cohorts across waves are independent by construction.
Conditional on $\F_{i-1}$, the increment of the product
$\prod_k R^{(k)}$ at any arrival is the increment of the single wave whose pair is
revealed, which has conditional mean at most one; the product is therefore itself a
nonnegative supermartingale started at one, for any adapted within-wave sampling times,
and Ville's inequality bounds its crossing probability by $\alpha$. Its power against
distributed leakage is additive in the exponent: if every wave carries advantage
$\Delta_0 > 0$, each factor drifts upward at rate
$\KL(\Bern(\tfrac12 + \tfrac{\Delta_0}2) \,\|\, \Bern(\tfrac12))$ per pair, so the
product crosses after roughly $\log(1/\alpha)/(m \cdot \KL)$ waves even when no single
wave is individually detectable.

\emph{(c)} Give wave $k$ its own alarm at level $\alpha_k = \alpha 2^{-(k+1)}$. By
Ville applied per wave, the probability that a truly clean wave $k$ ever alarms is at
most $\alpha_k$, and the union bound over an unbounded horizon costs
$\sum_{k \ge 1} \alpha 2^{-(k+1)} \le \alpha/2 \le \alpha$.
\end{proof}

\begin{proposition}[A magnitude-aware alarm]\label{prop:magnitude}
Under the null of Theorem~\ref{thm:null}, $\mathrm{sign}(D_i)$ is a fair coin even
conditional on $|D_i|$, so wealth factors of the form $1 + \lambda_i\,\mathrm{sign}(D_i)\,g_i$
are an exact e-process for any predictable stake $g_i \in [0,1]$ measurable with respect
to the past and $|D_i|$. Taking $g_i$ to be the predictable rank of $|D_i|$ concentrates
bets on the large-effect pairs where sparse memorisation lives, and halves detection
time under $5\%$-sparse leakage (Section~\ref{sec:exp}). Because it bets aggressively we
report it as an optional power enhancement and keep the sign-based arm as the default.
\end{proposition}

\begin{proof}
Symmetry of the conditional law of $D_i$ (Theorem~\ref{thm:null}) means that
conditioning on the magnitude $|D_i|$ leaves the sign exactly fair:
$\Efrak[\mathrm{sign}(D_i) \mid |D_i|, \F_{i-1}] = 0$, with ties randomised by the
committed coin. For any stake $g_i \in [0,1]$ that is measurable with respect to
$\F_{i-1} \vee \sigma(|D_i|)$---the predictable rank of $|D_i|$ among past magnitudes
qualifies---the tower property gives
$\Efrak[\mathrm{sign}(D_i)\, g_i \mid \F_{i-1}] = \Efrak[\,g_i\,
\Efrak[\mathrm{sign}(D_i) \mid |D_i|, \F_{i-1}] \mid \F_{i-1}] = 0$, so each factor
$1 + \lambda_i\, \mathrm{sign}(D_i)\, g_i$ with predictable $\lambda_i \in [0,1)$ has
conditional mean exactly one and stays positive. The product is a nonnegative
martingale with initial value one, hence an exact e-process, and Ville's inequality
applies as before. The power claim is empirical and quantified in
Section~\ref{sec:exp}.
\end{proof}

\begin{proposition}[Bridge from certified unlearning]\label{prop:bridge}
If $\Ucal$ is $(\eps_u,\delta_u)$-certified, then
$|\Delta^{(s)}| \le \tfrac{e^{\eps_u}-1}{e^{\eps_u}+1} + \delta_u$ for every $s$, so for
$\eps$ above this value VOUCH issues with probability $1 - \alpha - o(1)$. A certified
algorithm therefore passes VOUCH, as consistency demands, while VOUCH remains meaningful
at scales where $(\eps_u,\delta_u)$ certificates are unattainable.
\end{proposition}

\begin{proof}
The sign $Z^{(s)}$ is a randomised post-processing of the pipeline output: it is
computed from $M_u$ and the pair alone. $(\eps_u,\delta_u)$-certified unlearning makes
the law of $M_u$, under either value of the coin,
$(\eps_u,\delta_u)$-indistinguishable from the law of the retrained reference, under
which $Z^{(s)}$ is an exact fair coin by Theorem~\ref{thm:null}. Now view the rule
``guess that the twin with the larger score is the in-twin'' as a hypothesis test for
the coin $b_i$; its advantage is exactly
$\Pr[\text{correct}] - \Pr[\text{wrong}] = 2p^{(s)} - 1 = \Delta^{(s)}$. The
trade-off region of \citet{kairouz2015composition} states that any test between two
$(\eps_u,\delta_u)$-indistinguishable distributions has type-I and type-II errors
satisfying $e^{\eps_u}\beta_1 + \beta_2 \ge 1 - \delta_u$ and
$\beta_1 + e^{\eps_u}\beta_2 \ge 1 - \delta_u$; adding the two constraints,
$\beta_1 + \beta_2 \ge 2(1 - \delta_u)/(1 + e^{\eps_u})$, so the advantage obeys
\[
|\Delta^{(s)}| \;\le\; 1 - (\beta_1 + \beta_2)
\;\le\; 1 - \frac{2(1 - \delta_u)}{1 + e^{\eps_u}}
\;=\; \frac{e^{\eps_u} - 1}{e^{\eps_u} + 1} + \frac{2\delta_u}{1 + e^{\eps_u}}
\;\le\; \frac{e^{\eps_u} - 1}{e^{\eps_u} + 1} + \delta_u.
\]
When $\eps$ strictly exceeds this bound, every certificate process has strictly
positive drift by Theorem~\ref{thm:power} and crosses $1/\alpha$ with probability
tending to one as $m \to \infty$, while the revocation arm false-alarms with
probability at most $\alpha$; the issuance probability is therefore at least
$1 - \alpha - o(1)$.
\end{proof}

\paragraph{Scope, stated honestly.} A VOUCH certificate concerns influence extractable
through the declared class $\F$ on the canary law. It does not assert erasure in
parameter space, which no behavioural method can, and the extrapolation from canaries to
organic data is \emph{calibrated} through the dose--response strata rather than proven.
We regard stating these limits precisely as part of the contribution, since a
certificate whose scope is overclaimed is worse than none.

\section{Experiments}\label{sec:exp}
\input{sections/experiments}

\section{Limitations}\label{sec:limitations}

VOUCH certifies extractable influence through a declared score class on the canary law,
and three boundaries of that claim deserve emphasis. Parameter-space erasure is
behaviourally uncertifiable, so we do not claim it; the transfer from canaries to
organic data is calibrated by the dose--response strata, not proven; and the guarantee
is per-cohort rather than per-user, since per-request certificates would require
per-user canaries. The provider is assumed honest-but-verifiable, and a provider willing
to special-case the committed canaries needs cryptographic attestation that VOUCH does
not itself supply. Canaries must be planted before fine-tuning, so legacy models cannot
be retro-certified. Finally, both two-sided certification and consensus-based streaming
composition cost power, which we quantify (Tables~\ref{tab:power} and
\ref{tab:streaming}) so that cohorts can be sized accordingly rather than discovering
the cost after the fact.

\section{Conclusion}\label{sec:conclusion}

We set out to turn unlearning verification from a descriptive exercise into
certification, and the ingredients that made it possible were an exact null created by
the verifier's own randomness, a betting engine that survives continuous monitoring and
streaming deletions, and a pair of composition rules that we proved must run in opposite
directions. The empirical study shows the framework is not merely sound but practical:
it calibrates correctly under adversarial stopping, certifies genuine unlearning on
standard benchmarks across architectures, measures recoverability instead of arguing
about it, and does so at a verifier cost two orders of magnitude below shadow-model
evaluation. The released harness reproduces the entire study on free, ephemeral compute,
which we hope lowers the barrier to auditing unlearning wherever it is deployed.

\bibliography{refs}

\end{document}
