\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{parskip}
\usepackage{enumitem}
\pagestyle{empty}

\begin{document}

\begin{center}
{\Large\bfseries Point-by-Point Response to Editorial Comments}\\[6pt]
{\large VOUCH: Anytime-Valid Certificates for Machine Unlearning\\
via Paired-Canary Hypothesis-Betting}\\[4pt]
Manuscript ID: 4bbee415-ab8d-4923-a2b3-08f87b3b7b49\\[4pt]
Quang-Vinh Dang --- British University Vietnam\\[4pt]
16 July 2026
\end{center}

\vspace{6pt}
\noindent Dear Editors,

We thank the Editorial Office and the Editor-in-Chief for the technical check and for
identifying the missing Contribution Information Sheet. We have addressed both points
in full; our point-by-point response follows, with each editorial comment quoted
verbatim and our corresponding action described immediately below it.

\vspace{8pt}
\noindent\textbf{Comment 1 (Technical Check).}\\
\textit{``Please add the following required information on the first page of your
main manuscript file, right before the Abstract and keywords: A concise and
informative title; the full name(s) of the author(s); the affiliation(s) and
address(es) of the author(s) --- including the individual city and country
information; the e-mail address, telephone and fax numbers of the corresponding
author.''}

\emph{Response.} We have revised the first page of the manuscript accordingly. It
now presents, immediately before the Abstract:
\begin{itemize}[nosep,leftmargin=1.4em]
\item the full title, \emph{VOUCH: Anytime-Valid Certificates for Machine Unlearning
via Paired-Canary Hypothesis-Betting};
\item the full name of the (sole) author, Quang-Vinh Dang;
\item the affiliation and complete address, British University Vietnam, Ecopark,
Hung Yen, Vietnam;
\item the corresponding author's e-mail address, \texttt{vinh.dq4@buv.edu.vn}, and
telephone number, +84~768~496~935.
\end{itemize}
No fax number is provided, as the corresponding author does not have one; fax is not
in active use in our institution or country. We trust the e-mail and telephone
contact supplied satisfies the intent of this requirement.

\vspace{8pt}
\noindent\textbf{Comment 2 (Editor-in-Chief).}\\
\textit{``Your manuscript ... cannot currently be assessed because the mandatory
Contribution Information Sheet is missing ... Please kindly resubmit the paper
within 14 days with the CIS added in a separate file or as part of the cover
letter.''}

\emph{Response.} We have prepared the Contribution Information Sheet and included
it as part of the cover letter, immediately following the letter's closing, per the
option offered in this comment. It answers in full the four required questions: (i)
the main claim of the paper and its importance to the machine learning literature;
(ii) the theoretical, simulation, and end-to-end empirical evidence supporting the
claim, stated precisely; (iii) the papers by other authors that make the most
closely related contributions and how this paper relates to them; and (iv)
confirmation that no part of this paper has been published previously in any venue.

In preparing the sheet we also noted the guideline's remark that reviewers will
expect these answers to already be part of the paper's introductory material. On
review, the Introduction already stated the main claim and the three obstacles
VOUCH addresses (Comment 2, question~i) and Section~2 already gave a detailed,
explicit comparison against the closest related work (question~iii), but the
Introduction's discussion of supporting evidence (question~ii) described the
empirical study only qualitatively, without its headline quantitative results. We
have therefore added a paragraph to the Introduction (immediately before the
statement of contributions) that states these results directly --- the 2{,}000-seed
calibration outcomes, the measured size of both composition failures, detection and
certification rates across the benchmark study, the utility collapse under gradient
ascent, and the per-certificate computational cost --- so that this evidence is now
visible in the paper's own introductory material, consistent with the reviewers'
expectation and not only in the supplementary sheet.

\vspace{10pt}
\noindent We believe these revisions fully address both points raised, and we thank
the Editorial Office and the Editor-in-Chief for their guidance.

\vspace{14pt}
\noindent Sincerely,\\
Quang-Vinh Dang\\
British University Vietnam

\end{document}
