Our work sits at the meeting point of three literatures---unlearning algorithms,
the verification of unlearning, and game-theoretic statistics---and it is the gap
between the first two that game-theoretic statistics turns out to close. We survey
each in turn and then state precisely how VOUCH differs from its nearest neighbours.

\subsection{Unlearning algorithms and their benchmarks}

Exact unlearning by architectural design, exemplified by sharded training
\citep{bourtoule2021machine}, retrains only the shards touched by a deletion; it is
sound but cannot be applied to a model that was not trained for it, which is the
situation every deployed language model is in. The practical response has therefore
been \emph{approximate} unlearning, and for large language models a compact toolbox
has emerged: gradient ascent on the forget set and its retain-regularised variant
GradDiff; negative preference optimisation \citep{zhang2024npo}, which replaces the
unbounded ascent objective with a bounded preference loss and which its authors
introduced precisely to avoid the catastrophic collapse that ascent invites; the
simplified reappraisal SimNPO \citep{fan2025simnpo}; representation misdirection
\citep{li2024wmdp}; and task-vector negation. These methods are evaluated on
purpose-built benchmarks---TOFU's fictitious-author question answering
\citep{maini2024tofu}, MUSE's six-way news-and-books evaluation \citep{shi2024muse},
and the hazardous-knowledge multiple-choice of WMDP \citep{li2024wmdp}---and we treat
all of them, methods and benchmarks alike, as \emph{subjects} that VOUCH certifies
rather than as competitors.

What this literature has increasingly documented, however, is that the metrics used
to declare success are unreliable. A growing body of adversarial work shows that
apparent forgetting is often mere suppression:
\citet{lucki2024adversarial} recover ostensibly unlearned capabilities through
lightweight attacks; \citet{zhang2025quantization} show that simply quantising an
unlearned model to four bits resurrects the forgotten knowledge; \citet{deeb2024remove}
recover most of the pre-unlearning accuracy by fine-tuning on adjacent facts; and
\citet{fan2025relearning} formalise relearning attacks and connect robustness to
sharpness-aware minimisation. The evaluations themselves have come under equally
sharp scrutiny. \citet{feng2025existing} demonstrate that current LLM-unlearning
evaluations can be made to both overstate and understate success;
\citet{wei2025forget} show that facts presumed forgotten persist through correlated
knowledge; \citet{xu2026reversibility} distinguish genuine erasure from suppression
at the representation level and find that behaviour is easily restored by minimal
fine-tuning; and \citet{barbulescu2024each} document how per-sequence memorisation
strength governs how hard a datum is to remove. The recurring message across all of
this work is that \emph{evaluation}, not optimisation, is the bottleneck---which is
exactly the problem VOUCH addresses, and which motivates our recoverability probes
(R-VOUCH) and our insistence on pairing every certificate with a utility reading.

\subsection{Verifying and certifying unlearning}

The earliest guarantees came from \emph{certified removal}
\citep{guo2020certified,sekhari2021remember,neel2021descent,ginart2019making}, which
bounds the distance between the unlearned and retrained models in a differential-
privacy-style $(\eps_u,\delta_u)$ sense. These results are elegant but confined to
convex or near-convex regimes, and their bounds become vacuous at the scale of modern
language models; we use them not as a baseline but as a semantic anchor, showing in
Proposition~\ref{prop:bridge} that any $(\eps_u,\delta_u)$-certified algorithm passes
VOUCH. The verification of \emph{arbitrary} unlearning has proven far harder.
\citet{thudi2022necessity} argue that approximate unlearning cannot be audited from
weights alone and that auditable definitions are a prerequisite for any guarantee;
\citet{zhang2024verification} show that a dishonest provider can defeat both existing
families of verification strategies while passing them; and, most consequentially for
our design, \citet{zhang2024mia} establish that membership-inference scores
\emph{cannot} prove that a model was trained on given data unless the inclusion of
that data was randomised---the soundness principle that our fair-coin canary design
is built to satisfy. Two further impossibility results bound what any behavioural
certificate can claim: \citet{yu2025impossibility} show that path-dependence makes
retrain equivalence unattainable for multi-stage pipelines, which is why VOUCH's null
is internally generated rather than referenced to a retrained model, and the
reversibility findings above imply that no behavioural method can certify
parameter-space erasure---a limit we state explicitly rather than paper over.

A handful of recent efforts move toward statistical certification, and these are our
closest neighbours. \citet{pandey2025gaussian} give a hypothesis-testing notion of
certified unlearning, but it is fixed-sample and tied to a Gaussian high-dimensional
regression regime rather than to LLM-scale behavioural auditing.
\citet{pawelczyk2025machine} use data-poisoning as a lens on unlearning failure but
provide no error-controlled certificate. Cryptographic proof-of-unlearning
\citep{eisenhofer2025verifiable} attests that the correct computation \emph{ran},
which is orthogonal to whether forgetting \emph{holds} statistically and with which
VOUCH composes. The recent survey of \citet{xue2026survey} organises this fast-moving
area and confirms that no existing method delivers what VOUCH does: an anytime-valid,
distribution-free certificate of bounded residual influence. The broader
programme is surveyed by \citet{liu2025rethinking}.

\subsection{Game-theoretic statistics}

The inferential engine we borrow is that of safe, anytime-valid inference: e-values,
test martingales, and betting confidence sequences
\citep{ramdas2023game,grunwald2024safe,shafer2021testing,waudbysmith2024,howard2021time}.
The defining property, traceable to Ville's inequality \citep{ville1939}, is that a
nonnegative martingale started at one exceeds $1/\alpha$ with probability at most
$\alpha$ \emph{at every stopping time simultaneously}; this is precisely what a
streaming, continuously-monitored deletion audit requires, and precisely what a
fixed-sample test cannot offer. The betting perspective
\citep{waudbysmith2024,orabona2016coin} additionally supplies confidence sequences
whose bets adapt online to near-optimal growth, which we use both for the certificate
and for its quantitative confidence-sequence face. The one adjacent application of
these tools is to privacy auditing: \citet{steinke2023privacy} audit differential
privacy in a single training run, and \citet{gonzalez2025sequential} and
\citet{csillag2026dpevalues} bring sequential e-value machinery to DP auditing. The
distinction is fundamental to the hypothesis structure. Privacy auditing
\emph{lower}-bounds the privacy loss of a \emph{training} algorithm by mounting a
successful attack; VOUCH \emph{upper}-bounds the residual influence \emph{after
unlearning} by showing that no attack in a declared class succeeds---an equivalence
test, not a significance test, run over ghost twins that have passed through the
unlearning pipeline. To our knowledge, VOUCH is the first application of
game-theoretic statistics to unlearning certification.

\subsection{Positioning}

The contrast with our nearest neighbours can be summarised sharply. Fixed-sample
paired or two-sample tests---including the KS-based forget-quality metric of
\citet{maini2024tofu} and the min-$k\%$ leakage metric of \citet{shi2024muse}---are
invalidated by optional stopping and typically require a retrained reference; the
sound-membership results of \citet{zhang2024mia} tell us \emph{why} controlled
randomness is indispensable but stop at a fixed-$n$ membership proof rather than a
sequential forgetting certificate; shadow-model evaluators \citep{hayes2024inexact}
are descriptive and cost dozens of fine-tunes; certified removal
\citep{guo2020certified,sekhari2021remember} does not scale; and DP auditing
\citep{steinke2023privacy,gonzalez2025sequential} solves the opposite,
lower-bounding, problem. VOUCH is, to our knowledge, the first framework to formulate
unlearning certification as sequential equivalence testing and to deliver exact,
finite-sample, anytime-valid certificates of bounded residual influence at LLM scale,
with soundness inherited from a randomised paired-canary design rather than from
distributional assumptions or shadow models.
